---
title: "simpleRegress"
author: "Tiffany Alvarado"
date: "Fall 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```



the adjusted r-squared is 0.7476 for LENGTH ~ NECK.


## Model LENGTH ~ AGE this is an example

```{r}
simp <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(simp)
summary.lm(simp)
```

The adjusted r-squared for LENGTH ~ AGE is 0.5073.

## model LENGTH ~ HEADLEN this is an example

```{r}
simp <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(simp)
summary.lm(simp)
```
The adjusted rsquared for LENGTH ~ HEADLEN is 0.8434.

## Model LENGTH ~ HEADWTH this is an example


```{r}
simp <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(simp)
summary.lm(simp)
```
The rsquared for LENGTH ~ HEADWTH is 0.5321.

## Model LENGTH ~ CHEST this is an example


```{r}
simp <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(simp)
summary.lm(simp)
```
The adjusted rsquared for LENGTH ~ CHEST is 0.7869.

## Model LENGTH ~ WEIGHT this is an example

```{r}
simp <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(simp)
summary.lm(simp)
```
The adjusted rsquared for LENGTH ~ WEIGHT is 0.7423.



## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

All the models that predict Length from worst to best are:
LENGTH ~ AGE with the rsquare 0.5073.  
LENGTH ~ HEADWTH with the rsquared 0.5321.  
LENGTH ~ WEIGHT with the rsquared 0.7423.  
LENGTH ~ CHEST with the rsquared 0.7869.  
LENGTH HEADLEN with the rsquared 0.8434.  